Gesture Typing on Virtual Tabletop: Effect of Input Dimensions on Performance

### 2017-08-28 paper.tex
caption for touch classification
small edits

\section{Touch Classification}
This task is well suited to a supervised learning approach provided a training dataset is available. One of the authors video-taped 6 minutes of interactions creating 3 distincts datasets for 3 different fingers (index, thumb and pinky) totalling 10800 frames. The datasets are balanced with regards to touching frames and hovering frames and include a minor portion of out-of-range frames. The classification features are chosen as 20-bins histograms of z-values of the detected fingertip pointcloud, in red on~\autoref{fig:pointcloud}, which provides some orientation and shape invariance.

This is fed to a binary neural network classifier: 20 inputs dimensions, 2 layers of 64 units with Relu activation and $0.5$ dropout, and a fully connected layer with sigmoid activation. We use the binary cross-entropy as loss function and rmsprop for the optimiser. We train the model over 50 epochs and cross-validate with 3 folds (different fingers are thus validated against each others). The generalisation over different fingers is satisfying as shown by the average performance of 0.96 AUC for the ROC, see~\autoref{fig:roc_auc}. The model used for the user study was trained over the full dataset over 75 epochs. As an improvment, it is worth noting the possibility to delegate the feature extraction step to a model capable of inferring features directly, such as a convolutional neural network.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/roc_auc.pdf}
    \caption{Performance of the fingertip touch classification.}~\label{fig:roc_auc}
\end{figure}
}

@@

% =============================================================================
\section{Results}
% =============================================================================
The dependent variables we analyse are the success rate, the time taken per trial and the trace data when available. From these, we computed the dependent measure error rate defined as the percentage of unsucessful attempts as well as the text entry rate measured in Words Per Minute (WPM), as in~\cite{Markussen2014}, according to the following formula\footnote{$WPM = |T|/s \times 60/5$ where $|T|$ is the length of the transcribed string, s is time in seconds}.

@@

% We proposed a machine learning approach to fingertip touch detection. This problem is traditionally addressed by hand-tuning parameters and thresholding. In contrast, we showed that using a machine learning approach can solve the problem and generalise well enough to be used in an experiment with 12 unseen participants. One improvement could be to delegate the feature extraction step (an histogram of z-values in our case) to a model capable of inferring features directly, such as a convolutional neural network. Since the data is cheap to record, we could see this approach being used for many specific cases.

>>>>

\section{Touch Classification}
This task is well suited to a supervised learning approach provided a training dataset is available. One of the authors video-taped 6 minutes of interactions creating 3 distincts datasets for 3 different fingers (index, thumb and pinky) totalling 10800 frames. The datasets are balanced with regards to touching frames and hovering frames and include a minor fraction of out-of-range frames.

The classification features are chosen as 20-bins histograms of z-values of the detected fingertip pointcloud, in red on~\autoref{fig:pointcloud}, which provides some orientation and shape invariance. This is fed to a binary neural network classifier: 20 inputs dimensions, 2 layers of 64 hidden units with rectified linear (ReLU) activation and 50\% dropout, and a fully connected output layer with sigmoid activation. We use the binary cross-entropy as loss function and rmsprop for the optimiser. Finally, we train the model over 50 epochs and perform a 3 folds cross-validation across the 3 datasets, different fingers are thus validated against each others. This provides an estimate for the generalisation power of the model when classifying unseen users as shown by the averaged 0.96 AUC for the ROC, see~\autoref{fig:roc_auc}.

The model used for the user study was trained over the full dataset over 75 epochs. As an improvment, it is worth noting the possibility to delegate the feature extraction step to a model capable of inferring features directly, such as a convolutional neural network.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/roc_auc.pdf}
    \caption{Performance of the fingertip touch classification with 3-fold cross validation. Each split trains on two finger types (among three) and validates against the remaining one.}~\label{fig:roc_auc}
\end{figure}
}

@@

% =============================================================================
\section{Results}
% =============================================================================
The dependent variables we analyse are the success rate, the time taken per trial and the trace data when available. From these, we computed the dependent measure error rate defined as the percentage of unsucessful attempts as well as the text entry rate measured in Words Per Minute (WPM), as in~\cite{Markussen2014}, according to the following formula: $WPM = |T|/s \times 60/5$ where $|T|$ is the length of the transcribed string and $s$ is time in seconds.

@@
